# Run the whole pipeline

> One File to rule them all, 
> 
> One File to find them (in online databases),
> 
> One File to bring them all (into R), 
> 
> and in the darkness bind them (into a cleaned csv)
> - J.R.R. Tolkien (probably)

This file will run the entire pipeline, for all sources. Comment out / don't run blocks if you don't need them.

Please check the `last_successful_run.md` file after running this, to see if all files were successfully created and saved (make sure to take a look at the Status column in the Table).

All in all, this code should not take much longer than 5 minutes

## Utilities

```{r}
library(quarto)
```

Dedicated Runner to catch errors and clean up afterwards (quarto creates a lot of unneeded files running a QMD file)
```{r}
my_quarto_render <- function(source, file_name) {
  tryCatch({ # Try to run the code
    print(paste0("Running: ", file_name))
    # run the code
    quarto_render(
      input = paste0("./src/", source,"/",file_name, ".qmd"),
      output_file = "NULL",
      quiet = T
    )
    # remove the output file 
    unlink(paste0("./src/", source,"/", "NULL.html"))
    # remove the associated folder
    unlink(paste0("./src/", source,"/", file_name, "_files"), recursive = TRUE, force = TRUE)
  }, error = function(e) { # Catch any errors during running
    print(paste0("ERROR: the code did not run, take a look at the file and fix it from there"))
  }
  )
}
```

## IMF

```{r}
my_quarto_render("imf", "02_monthly_data")
my_quarto_render("imf", "03_annual_data")
```

## World Bank

```{r}
my_quarto_render("wb", "01_wdi_download")
my_quarto_render("wb", "02_WEO_proxy_GGEI")
```

## Eurostat

Note: Make sure to run all of them at once, so that the combined outputted file is up-to-date! 
```{r}
my_quarto_render("eurostat", "02_national_accounts") # National Accounts
my_quarto_render("eurostat", "03_CPI") # Consumer Price Index
my_quarto_render("eurostat", "04_Unempl") # Unemployment

my_quarto_render("eurostat", "05_File_combination") # Combine the files and save as one file with multiple sheets
```

## Le fin

Hopefully it all worked. If not, get into the weeds and take a look at the individual files to see where the error is. Youll probably have to adjust some links, because there was an internal update in the databases or similar.
