# WDI Indicator Download

Required Indicators

| Names                                                        | WEFM Code | WDI Code          |
| ------------------------------------------------------------ | --------- | ----------------- |
| Interest payments (current LCU)                              | GGEI      | GC.XPN.INTP.CN    |
| GNI per capita, Atlas method (current US$)                   | GNICAP    | NY.GNP.PCAP.CD    |
| Gini index                                                   | GINI      | SI.POV.GINI       |
| Survey mean consumption or income per capita, total population (2021 PPP $ per day) | YBAR      | SI.SPR.PCAP       |
| Poverty headcount ratio at $3.00 a day (2021 PPP) (% of population) | HEAD      | SI.POV.DDAY       |
| PPP conversion factor, GDP (LCU per international $)         |           | PA.NUS.PPP        |
| GDP, PPP (current international $)                           |           | NY.GDP.MKTP.PP.CD |
| GDP, PPP (constant 2021 international $)                     |           | NY.GDP.MKTP.PP.KD |

## Load

```{r}
library(WDI)
library(tidyverse)
```

```{r}
df_wdi_raw <- WDI(
  country = "all",
  indicator = c(
    "GNICAP" =  "NY.GNP.PCAP.CD",
    "GINI" = "SI.POV.GINI",
    "GGEI" = "GC.XPN.INTP.CN",
    "YBAR" = "SI.SPR.PCAP",
    "HEAD" = "SI.POV.DDAY",
    "PPP" = "PA.NUS.PPP",
    "GDP_PPP" = "NY.GDP.MKTP.PP.CD",
    "GDP_PPP_KD" = "NY.GDP.MKTP.PP.KD"
  ),
  start = 1970,
  extra = T
)
```

## Clean

Clean the dataset and only include countries in WEFM

```{r}
df_wesp <- readxl::read_excel("../../data/raw/Regions.xlsx")

wesp_countries <- unique(df_wesp$ISO)
```


```{r}
df_wdi <- df_wdi_raw %>%
  filter(
    region != "Aggregates",
    # iso3c %in% c("DEU", "GBR")
    iso3c %in% wesp_countries
    ) %>%
  select(iso3c, year, GNICAP, GINI, GGEI, YBAR, HEAD, PPP, GDP_PPP, GDP_PPP_KD) %>%
  arrange(iso3c, year) %>%
  pivot_wider(
    names_from = iso3c, 
    values_from = c(GNICAP, GINI, GGEI, YBAR, HEAD, PPP, GDP_PPP, GDP_PPP_KD), 
    names_glue = "{iso3c}_{.value}"
    ) %>%
  # filter out columns with all NA values
  select(where(~ !all(is.na(.))))

df_wdi
```



## Save

In case you want to save it as a simple csv, instead of calling the whole timestamping process
```{r}
# write_csv(df_wdi, "../../data/wb_processed/wdi_indicators.csv")
```

But for production, use the timestamping function to save some metadata
```{r}
source("../utils/save_timestamp.R")

timestamp_save(
  df = df_wdi, 
  filename = "wdi_indicators", 
  source = "WB", 
  frequency = "annual", 
  format = "csv"
)
```

