# Comparison of GGEI data

Three sources

- IMF IFS
- WB WDI
- IMF WEO Proxy, from:
  - GGXOFB:	General government overall fiscal balance
  - GGXONLB	General government primary net lending/borrowing

```{r}
library(tidyverse)
```

## Data

### WDI
```{r}
df_wdi_raw <- read_csv("../../data/wb_processed/wdi_indicators.csv")
```

```{r}
df_wdi_GGEI <- df_wdi_raw %>%
  select(year, contains("GGEI")) %>%
  pivot_longer(cols = contains("GGEI"), names_to = "country", values_to = "GGEI") %>%
  mutate(country = str_replace(country, "_GGEI", ""))

df_wdi_GGEI
```

### WEO data

```{r}
df_weo_raw <- read_csv("../../data/wb_processed/wdi_indicators_with_proxy.csv")
```

```{r}
df_weo <- df_weo_raw %>%
  select(year, contains("WEO_proxy")) %>%
  pivot_longer(cols = contains("WEO_proxy"), names_to = "country", values_to = "GGEI_proxy") %>%
  mutate(country = str_replace(country, "_WEO_proxy", ""))
```

### IMF old data:
```{r}
df_old_adv <- read_csv("../../examples/imf/adv_annual.csv")
df_old_dev <- read_csv("../../examples/imf/developing_annual.csv")
df_old_oth <- read_csv("../../examples/imf/other_annual.csv")

imf_CONV <- readxl::read_excel("../../data/raw/imf_CONV.xlsx") 
```

```{r}
df_old_comb <- tibble()
for (df in list(df_old_adv, df_old_dev, df_old_oth)) {
  df <- df %>%
    select(`Country Name`, `Country Code`, `Indicator Name`, `Indicator Code`, Attribute, `1960`:`2024`) %>%
    mutate(across(`1960`:`2024`, as.numeric))
  df_old_comb <- bind_rows(df_old_comb, df) 
}
rm(df_old_adv, df_old_dev, df_old_oth)
```

```{r}
df_old <- df_old_comb %>%
  filter(`Indicator Code` == "GG_GEI_G01_XDC") %>%
  mutate(
    `Country Code` = as.character(`Country Code`),
    ) %>%
  left_join(imf_CONV, by = c("Country Code" = "Code")) %>%
  select(ISO, `1960`:`2024`) %>%
  pivot_longer(cols = `1960`:`2024`, names_to = "year", values_to = "IFS_GGEI") %>%
  mutate(year = as.numeric(year))
```



## Combine

```{r}
df_wesp <- readxl::read_excel("../../data/raw/Regions.xlsx")

wesp_countries <- unique(df_wesp$ISO)
```

```{r}
df_comb <- df_weo %>%
  filter(year >= 2000) %>%
  left_join(df_wdi_GGEI %>% filter(year >= 2000), by = c("country", "year")) %>%
  left_join(df_old %>% filter(year >= 2000), by = c("country" = "ISO", "year")) %>%
  filter(country %in% wesp_countries)
```

## Comparison

Messy code, but it creates some comparison tables for
- number of countries with any data
- number of countries with more than 10 years of data
- number of countries with full data between 2000 and 2010 (WEFM training period)

```{r}
df_any_data <- df_comb %>%
  group_by(country) %>%
  reframe(
    n_years_weo = sum(!is.na(GGEI_proxy)),
    n_years_wdi = sum(!is.na(GGEI)),
    n_years_ifs = sum(!is.na(IFS_GGEI))
  ) %>%
  summarise(
    stat = "Nr of countries with any data",
    weo = sum(n_years_weo > 0),
    wdi = sum(n_years_wdi > 0),
    ifs = sum(n_years_ifs > 0)
  ) 

df_more_than10 <- df_comb %>%
  group_by(country) %>%
  reframe(
    n_years_weo = sum(!is.na(GGEI_proxy)),
    n_years_wdi = sum(!is.na(GGEI)),
    n_years_ifs = sum(!is.na(IFS_GGEI))
  ) %>%
  summarise(
    stat = "Nr of countries with more than 10 years of data",
    weo = sum(n_years_weo > 10),
    wdi = sum(n_years_wdi > 10),
    ifs = sum(n_years_ifs > 10)
  )

df_2000_2010 <- df_comb %>%
  filter(year >= 2000, year <= 2010) %>%
  group_by(country) %>%
  reframe(
    n_years_weo = sum(!is.na(GGEI_proxy)),
    n_years_wdi = sum(!is.na(GGEI)),
    n_years_ifs = sum(!is.na(IFS_GGEI))
  ) %>%
  summarise(
    stat = "Nr of countries with full data between 2000 and 2010",
    weo = sum(n_years_weo > 10),
    wdi = sum(n_years_wdi > 10),
    ifs = sum(n_years_ifs > 10)
  )

bind_rows(df_any_data, df_more_than10, df_2000_2010)
```

Correlation matrix
```{r}
df_comb %>%
  select(GGEI_proxy, GGEI, IFS_GGEI) %>%
  rename(
    WEO = GGEI_proxy,
    WDI = GGEI,
    IFS = IFS_GGEI
  ) %>%
  cor(use = "pairwise.complete.obs")
```

Plot for one country
```{r}
df_comb %>%
  filter(country == "GBR") %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y = GGEI_proxy, color = "WEO Proxy")) +
  geom_line(aes(y = GGEI, color = "WDI GGEI")) +
  geom_line(aes(y = IFS_GGEI, color = "IMF IFS GGEI")) +
  facet_wrap(~ country) + # in case I want to do this for multiple countries
  labs(title = "Comparison of GGEI from WEO Proxy, WDI, and IMF IFS",
       x = "Year",
       y = "Value",
       color = "Data Source") 
```