# Looking at the Eurostat Data

Sources:

- National Accounts data: 
  - annual: [nama_10_gdp](https://ec.europa.eu/eurostat/databrowser/view/nama_10_gdp/default/table?lang=en)
  - quarterly: [namq_10_gdp](https://ec.europa.eu/eurostat/databrowser/view/namq_10_gdp/default/table?lang=en)
- Inflation data: 
  - monthly: [prc_hicp_midx](https://ec.europa.eu/eurostat/databrowser/view/prc_hicp_midx/default/table?lang=en)
  - annual: [prc_hicp_aind](https://ec.europa.eu/eurostat/databrowser/view/prc_hicp_aind/default/table?lang=en)
- Unemployment data
  - monthly: [une_rt_m](https://ec.europa.eu/eurostat/databrowser/view/une_rt_m/default/table?lang=en)
  - annual: [une_rt_a](https://ec.europa.eu/eurostat/databrowser/view/une_rt_a/default/table?lang=en)

## Loading

```{r}
library(tidyverse)
library(readxl)
```

The excel sheet is a mess:
```{r}
excel_sheets("../../examples/eurostat/Eurostat.xlsx")
```

## Annual

```{r}
df_annual <- read_excel("../../examples/eurostat/Eurostat.xlsx", sheet="Annual",)
df_annual <- df_annual %>%
  rename(
    `indicator` = `...1`,
    `country` = `...2`,
    `unit` = `...3`,
    `code` = `...4`
  )
```

```{r}
df_annual %>%
  filter(country=="Germany") %>%
  select(indicator, unit, code) %>%
  unique()
```

Vars: Current Prices + Chain linked volumnes (2010=100): - GDP at market prices - Final consumption expenditure of general government - Household and NPISH final consumption expenditure - Gross fixed capital formation - Exports of goods and services - Imports of goods and services

-   All-items HICP: Annual Average Index
-   Unemployment Rate: Total

| Desc | Unit | Code |
|----|----|----|
| GDP at market prices | Current + Chain | XXX_YER_ESTAT   |
| Final consumption expenditure of general government | \-""- | XXX_GCR_ESTAT |
| Household and NPISH final consumption expenditure | \-""- | XXX_PCR_ESTAT |
| Gross fixed capital formation | \-""- | XXX_ITR_ESTAT |
| Exports of goods and services | \-""- | XXX_XTR_ESTAT |
| Imports of goods and services | \-""- | XXX_MTR_ESTAT |
| All-items HICP | Annual Average Index | XXX_HIC_ESTAT |
| Unemployment Rate | Total | XXX_URX_ESTAT |

## Quarterly

```{r}
df_quarterly <- read_excel("../../examples/eurostat/Eurostat.xlsx", sheet="Quarterly",)
```

```{r}
df_quarterly <- df_quarterly %>%
  rename(
    indicator = `...1`,
    country = `...2`,
    unit = `...3`,
    adjustment = `...4`,
    code = `...5`
  )
```

```{r}
df_quarterly %>%
  filter(country=="Germany") %>%
  select(indicator,adjustment, unit) %>%
  unique()
```

- Same 6 macro variables
- by chain lined volumes+ current prices
- by adjustment: seasonally+not calendar,seasonal+calendar,unadjusted 
= 6x3x2 = 36 variables

