# National Accounts

Table: [nama_10_gdp](https://ec.europa.eu/eurostat/databrowser/view/nama_10_gdp/default/table?lang=en)

Countries: all
Years: 2010-current

ESA Indicators:
- B1GQ: GDP at market prices
- P3_S13: Final consumption expenditure of general government
- P31_S14_S15: Household and NPISH final consumption expenditure
- P51G: Gross fixed capital formation
- P6: Exports of goods and services
- P7: Imports of goods and services

Units:
- CLV_I10: Chain linked volumes (2010=100)
- CP_MNAC: Current prices, million euro

for Quarterly: Adjusmtnets
- NSA: Not seasonally adjusted
- SA: Seasonally adjusted
- SCA: Seasonally adjusted and calendar adjusted


```{r}
library(tidyverse)
library(eurostat)
library(countrycode)
```

Sometimes the Eurostat cache can cause issues, especially if there have been updates to the data. If you encounter any problems with the data retrieval, you can try clearing the cache using the following command:
```{r}
#clean_eurostat_cache()
```


## Annual national accounts

```{r}
df_annual_raw <- get_eurostat(
  "nama_10_gdp", 
  filters=list(
    na_item = c("B1GQ", "P3_S13", "P31_S14_S15", "P51G", "P6", "P7"),
    unit = c("CLV_I10", "CP_MNAC"),
    sinceTimePeriod=2000
    ),
  )

df_annual_raw <- label_eurostat(df_annual_raw, code=c("na_item","unit"), countrycode=c("iso3c") )
```

```{r}
df_annual <- df_annual_raw %>%
  filter(str_length(geo) == 3) %>% # only countries, no aggregates
  mutate(
    WEFM_code = case_when(
      # chain linked codes
      na_item_code == "B1GQ" & unit_code == "CLV_I10" ~ "YER",
      na_item_code == "P3_S13" & unit_code == "CLV_I10" ~ "GCR",
      na_item_code == "P31_S14_S15" & unit_code == "CLV_I10" ~ "PCR",
      na_item_code == "P51G" & unit_code == "CLV_I10" ~ "ITR",
      na_item_code == "P6" & unit_code == "CLV_I10" ~ "XTR",
      na_item_code == "P7" & unit_code == "CLV_I10" ~ "MTR",
      # current price codes
      na_item_code == "B1GQ" & unit_code == "CP_MNAC" ~ "YEN",
      na_item_code == "P3_S13" & unit_code == "CP_MNAC" ~ "GCN",
      na_item_code == "P31_S14_S15" & unit_code == "CP_MNAC" ~ "PCR",
      na_item_code == "P51G" & unit_code == "CP_MNAC" ~ "ITN",
      na_item_code == "P6" & unit_code == "CP_MNAC" ~ "XTN",
      na_item_code == "P7" & unit_code == "CP_MNAC" ~ "MTN",
    ),
    FINAL_CODE = paste0(geo, "_", WEFM_code, "_ESTAT"),
    year = year(time),
    country_name = countrycode(geo, "iso3c", "country.name"),
    indicator = na_item
  ) %>%
  select(indicator, country_name, unit, FINAL_CODE, year, values) %>%
  pivot_wider(names_from = year, values_from = values) 
```

```{r}
df_annual
```

```{r}
write_csv(df_annual, "../../data/eurostat_processed/National_Accounts_A.csv")
```


## Quarterly national accounts

```{r}
df_quarterly_raw <- get_eurostat(
  "namq_10_gdp", 
  filters=list(
    na_item = c("B1GQ", "P3_S13", "P31_S14_S15", "P51G", "P6", "P7"),
    unit = c("CLV_I10", "CP_MNAC"),
    s_adj = c("NSA", "SA", "SCA"),
    sinceTimePeriod=2000
  ),
)

df_quarterly_raw <- label_eurostat(df_quarterly_raw, code=c("na_item","unit","s_adj"), countrycode=c("iso3c") )
```

```{r}
df_quarterly <- df_quarterly_raw %>%
  filter(str_length(geo) == 3) %>%# only countries, no aggregates
  mutate(
    WEFM_code = case_when(
      # chain linked codes
      na_item_code == "B1GQ" & unit_code == "CLV_I10" ~ "YER",
      na_item_code == "P3_S13" & unit_code == "CLV_I10" ~ "GCR",
      na_item_code == "P31_S14_S15" & unit_code == "CLV_I10" ~ "PCR",
      na_item_code == "P51G" & unit_code == "CLV_I10" ~ "ITR",
      na_item_code == "P6" & unit_code == "CLV_I10" ~ "XTR",
      na_item_code == "P7" & unit_code == "CLV_I10" ~ "MTR",
      # current price codes
      na_item_code == "B1GQ" & unit_code == "CP_MNAC" ~ "YEN",
      na_item_code == "P3_S13" & unit_code == "CP_MNAC" ~ "GCN",
      na_item_code == "P31_S14_S15" & unit_code == "CP_MNAC" ~ "PCR",
      na_item_code == "P51G" & unit_code == "CP_MNAC" ~ "ITN",
      na_item_code == "P6" & unit_code == "CP_MNAC" ~ "XTN",
      na_item_code == "P7" & unit_code == "CP_MNAC" ~ "MTN",
    ),
    FINAL_CODE = paste0(geo, "_", WEFM_code, "_",s_adj_code, "_ESTAT_Q"),
    year = year(time),
    quarter = quarter(time),
    time_period = paste0(year, "Q", quarter),
    country_name = countrycode(geo, "iso3c", "country.name"),
    indicator = na_item
  ) %>%
  select(indicator, country_name, unit, s_adj, FINAL_CODE,time_period, values) %>%
  pivot_wider(names_from = time_period, values_from = values)

df_quarterly
```

```{r}
write_csv(df_quarterly, "../../data/eurostat_processed/National_Accounts_Q.csv")
```

