# Combining the tables

into one excel file, with all the quirks that come with it

```{r}
library(writexl)
library(tidyverse)
```

## Read in the processed data

```{r message=FALSE}
df_ANA <- read_csv("../../data/eurostat_processed/National_Accounts_A.csv")
df_ANQ <- read_csv("../../data/eurostat_processed/National_Accounts_Q.csv")

df_CPIA <- read_csv("../../data/eurostat_processed/CPI_A.csv")
df_CPIM <- read_csv("../../data/eurostat_processed/CPI_M.csv")

df_UR_A <- read_csv("../../data/eurostat_processed/UR_A.csv")
df_UR_M <- read_csv("../../data/eurostat_processed/UR_M.csv")
```

**Build the annual sheet**
```{r}
df_annual <- df_ANA %>%
  bind_rows(df_CPIA) %>%
  bind_rows(df_UR_A) 
```


**Build the monthly sheet**

Sometimes, the dimensions dont add up, because one dataset already has the most recent month, while the other one doesnt
```{r}
if (nrow(df_CPIM) != nrow(df_UR_M)) {
  # catch the difference in the date columns
  intersection_dates <- intersect(df_CPIM$time, df_UR_M$time)
  df_CPIM <- df_CPIM %>% filter(time %in% intersection_dates)
  df_UR_M <- df_UR_M %>% filter(time %in% intersection_dates)
}
```


```{r}
df_monthly <- df_CPIM %>%
  # merge horizontally with the UR_M table, which has the same structure
  bind_cols(df_UR_M %>% select(-time), .name_repair = "unique") 
```


## Filter UK out

Due to Brexit, UK data is messy, and has potential problems in the future.

```{r}
df_annual <- df_annual %>% filter(country_name != "United Kingdom")

df_monthly <- df_monthly %>% select(-c(GBR_URX_ESTAT_M, GBR_HIC_ESTAT_M)) 

df_ANQ <- df_ANQ %>% filter(country_name != "United Kingdom")
```

## Write to excel

```{r}
source("../utils/save_timestamp.R")

timestamp_save(
  filename = "EUROSTAT",
  source = "EUROSTAT",
  frequency = "Mixed",
  format = "xlsx",
  df = list(
    Annual = df_annual, 
    Quarterly = df_ANQ,
    Monthly = df_monthly
  )
)
```

