# Statistics 

comparing the old and the new one (especially the annual data)

## Loading

```{r}
#| message: false
library(tidyverse)
library(visdat)
library(patchwork)
library(skimr)
```

```{r}
#| message: false
#| warning: false
df_new <- read_csv("../../data/imf_processed/all_annual.csv")

df_old_adv <- read_csv("../../examples/imf/adv_annual.csv")
df_old_dev <- read_csv("../../examples/imf/developing_annual.csv")
df_old_oth <- read_csv("../../examples/imf/other_annual.csv")
```

Combine the old ones
```{r}
df_old_comb <- tibble()
for (df in list(df_old_adv, df_old_dev, df_old_oth)) {
  df <- df %>%
    select(`Country Name`, `Country Code`, `Indicator Name`, `Indicator Code`, Attribute, `1960`:`2024`) %>%
    mutate(across(`1960`:`2024`, as.numeric))
  df_old_comb <- bind_rows(df_old_comb, df) 
}
rm(df_old_adv, df_old_dev, df_old_oth)
```

```{r}
df_new <- df_new %>%
  select(`Country Name`, `Country Code`, `Indicator Name`, `Indicator Code`, Attribute, `1960`:`2024`) %>%
  mutate(across(`1960`:`2024`, as.numeric)) %>%
  arrange(`Indicator Code`, `Country Code`)
```

Filtering the old one for the new indicators
```{r}
df_old <- df_old_comb %>%
  filter(
    `Indicator Code` %in% unique(df_new$`Indicator Code`),
    Attribute == "Value"
    ) %>%
  arrange(`Indicator Code`, `Country Code`) %>%
  mutate(`Country Code` = as.character(`Country Code`))
```

## Ex.: ENDA_XDC_USD_RATE

```{r}
indicator <- "ENDA_XDC_USD_RATE"
df_new_ind <- df_new %>%
  filter(`Indicator Code` == indicator)
df_old_ind <- df_old %>%
  filter(`Indicator Code` == indicator)
```

Lets start with a visual comparison
```{r}
vis_dat(df_old_ind) + vis_dat(df_new_ind) +
  plot_annotation(title = indicator) +
  plot_layout(guides = "collect")
```
=> Looks very similar in structure

Differences in country codes present in new, but not in old
```{r}
present_new <- setdiff(df_new_ind$`Country Code`, df_old_ind$`Country Code`)
df_new_ind %>%
  filter(`Country Code` %in% present_new) %>%
  select(`Country Name`, `Country Code`) %>%
  distinct()
```
= well USSR is quote outdated, and Caymen Islands is not that important

Other way around
```{r}
present_old <- setdiff(df_old_ind$`Country Code`, df_new_ind$`Country Code`)
df_old_ind %>%
  filter(`Country Code` %in% present_old) %>%
  select(`Country Name`, `Country Code`) %>%
  distinct()
```
= seems to be mostly small islands belinging to the UK, regional groupings, and an old code for North yemen (existed until 1990)

Lets filter them out
```{r}
df_old_ind <- df_old_ind %>% filter(!`Country Code` %in% present_old) 
df_new_ind <- df_new_ind %>% filter(!`Country Code` %in% present_new)
```

Now we can compare the actual values
```{r}
compare <- diffdf::diffdf(
  df_old_ind, df_new_ind,
  strict_numeric = F,
  keys = c("Country Code"),
  tolerance = 0.001
)
```

```{r}
compare$NumDiff
# compare                # print all of it
# compare$VarDiff_2017   # example of differences in 2017, the old one does not have data for it...
```

- Some names differ for countries (as expected)
- the description has a typo (that should not be relevant)
- Values are almost identical (just in some years there is one or two observations missing)

For quick comparison, lets build some summary statistics 

```{r}
skim_df1 <- skim(df_new_ind) %>% reframe(skim_variable, numeric.mean, n_missing,source = "new")
skim_df2 <- skim(df_old_ind) %>% reframe(skim_variable, numeric.mean, n_missing,source = "old")

skim_df1 %>% 
  bind_rows(skim_df2) %>%
  arrange(skim_variable, source)
```

## Functions

Automate the analysis

```{r}
get_dfs <- function(indicator) {
  df_new_ind <- df_new %>%
    filter(`Indicator Code` == indicator)
  df_old_ind <- df_old %>%
    filter(`Indicator Code` == indicator)
  list(new = df_new_ind, old = df_old_ind)
}
```

```{r}
visualize <- function(df_old_ind, df_new_ind, indicator) {
  vis_dat(df_old_ind) + vis_dat(df_new_ind) +
    plot_annotation(title = indicator) +
    plot_layout(guides = "collect")
}
```

```{r}
get_country_diffs <- function(df_old_ind, df_new_ind) {
  present_new <- setdiff(df_new_ind$`Country Code`, df_old_ind$`Country Code`)
  present_old <- setdiff(df_old_ind$`Country Code`, df_new_ind$`Country Code`)
  # print codes and names in list

  df_old_ind %>%
    filter(`Country Code` %in% present_old) %>%
    select(`Country Name`, `Country Code`) %>%
    rename(`Only in Old` = "Country Name") %>%
    distinct() %>%
    print()
  df_new_ind %>%
    filter(`Country Code` %in% present_new) %>%
    select(`Country Name`, `Country Code`) %>%
    rename(`Only in New` = "Country Name") %>%
    distinct() %>%
    print()
}
```

```{r}
filter_diffs <- function(df_old_ind, df_new_ind) {
  present_new <- setdiff(df_new_ind$`Country Code`, df_old_ind$`Country Code`)
  present_old <- setdiff(df_old_ind$`Country Code`, df_new_ind$`Country Code`)
  
  df_old_ind <- df_old_ind %>% filter(!`Country Code` %in% present_old) 
  df_new_ind <- df_new_ind %>% filter(!`Country Code` %in% present_new)
  list(old = df_old_ind, new = df_new_ind)
}
```

```{r}
compare_values <- function(df_old_ind, df_new_ind, tolerance = 0.001) {
  compare <- diffdf::diffdf(
    df_old_ind, df_new_ind,
    strict_numeric = F,
    keys = c("Country Code"),
    tolerance = tolerance
  )
  compare
}
```

```{r}
plot_compare <- function(compare, indicator) {
  compare$NumDiff %>% 
    filter(Variable != "Country Name") %>%
    ggplot(aes(x = Variable, y = `No of Differences`)) +
    geom_col() +
    plot_annotation(
      title = paste("Number of differences for", indicator),
      subtitle = paste("Total countries", compare$DataSummary$BASE[[3]])
      )
}
```

```{r}
summary_statistics <- function(df_new_ind, df_old_ind, compare) {
  pct_diff <- sum( # nr of differences in values divided by total number of cells
    compare$NumDiff %>% filter(Variable != "Country Name") %>% pull(`No of Differences`)) /
    (as.numeric(compare$DataSummary$BASE[[3]]) *   # rows
       as.numeric(compare$DataSummary$BASE[[4]])   # columns
     )

  df_new_ind %>%
    reframe(
      Indicator_Code = first(`Indicator Code`),
      Indicator_Name = str_trunc(first(`Indicator Name`), 30, "right"),
      n_countries = n_distinct(`Country Code`),
      pct_NA = mean(is.na(across(`1960`:`2024`) %>% as.matrix())),
      pct_NA_2010_2020 = mean(is.na(across(`2010`:`2020`) %>% as.matrix())),
      pct_diff = pct_diff
    )
}
```


## PCPI_PC_CP_A_PT

```{r}
dfs_pcpi <- get_dfs("PCPI_PC_CP_A_PT")
visualize(dfs_pcpi$old, dfs_pcpi$new, "PCPI_PC_CP_A_PT")
```

```{r}
get_country_diffs(dfs_pcpi$old, dfs_pcpi$new)
```
similar to before, some regional groupings and some small islands

```{r}
filtered_dfs_pcpi <- filter_diffs(dfs_pcpi$old, dfs_pcpi$new)
compare_pcpi <- compare_values(filtered_dfs_pcpi$old, filtered_dfs_pcpi$new, tolerance = 0.01)
```

```{r}
plot_compare(compare_pcpi, "PCPI_PC_CP_A_PT")
```
in recent years = much more differences, probably data updates

```{r}
summary_pcpi <- summary_statistics(filtered_dfs_pcpi$new, filtered_dfs_pcpi$old, compare_pcpi)
summary_pcpi
```

## PCPI_IX

```{r}
dfs_pcpi_ix <- get_dfs("PCPI_IX")
visualize(dfs_pcpi_ix$old, dfs_pcpi_ix$new, "PCPI_IX")
```

```{r}
get_country_diffs(dfs_pcpi_ix$old, dfs_pcpi_ix$new)
```
same countries / regions as before

```{r}
filtered_dfs_pcpi_ix <- filter_diffs(dfs_pcpi_ix$old, dfs_pcpi_ix$new)
compare_pcpi_ix <- compare_values(filtered_dfs_pcpi_ix$old, filtered_dfs_pcpi_ix$new, tolerance = 0.0001)
```
```{r}
plot_compare(compare_pcpi_ix, "PCPI_IX")
```
later years = more differences, even more than before

```{r}
summary_pcpi_ix <- summary_statistics(filtered_dfs_pcpi_ix$new, filtered_dfs_pcpi_ix$old, compare_pcpi_ix)
summary_pcpi_ix
```

## BCAXF_BP6_USD

```{r}
dfs_bcaxf <- get_dfs("BCAXF_BP6_USD")
visualize(dfs_bcaxf$old, dfs_bcaxf$new, "BCAXF_BP6_USD")
```
that looks really good already

```{r}
get_country_diffs(dfs_bcaxf$old, dfs_bcaxf$new)
```
The new data has a lot more countries

```{r}
filtered_dfs_bcaxf <- filter_diffs(dfs_bcaxf$old, dfs_bcaxf$new)
compare_bcaxf <- compare_values(filtered_dfs_bcaxf$old, filtered_dfs_bcaxf$new, tolerance = 1e+6)
```

```{r}
plot_compare(compare_bcaxf, "BCAXF_BP6_USD")
```
and also a lot more data in recent years

```{r}
summary_bcaxf <- summary_statistics(filtered_dfs_bcaxf$new, filtered_dfs_bcaxf$old, compare_bcaxf)
summary_bcaxf
```

If we take a closer look at the differences, it is quite clear that they are small and mostly data updates
```{r}
compare_bcaxf$VarDiff_2017
```

## FMB_XDC

```{r}
dfs_fmb <- get_dfs("FMB_XDC")
visualize(dfs_fmb$old, dfs_fmb$new, "FMB_XDC")
```
Note that the new one also has much more countries, looks less than it is

```{r}
get_country_diffs(dfs_fmb$old, dfs_fmb$new)
```
```{r}
filtered_dfs_fmb <- filter_diffs(dfs_fmb$old, dfs_fmb$new)
compare_fmb <- compare_values(filtered_dfs_fmb$old, filtered_dfs_fmb$new, tolerance = 1e+12)
```
```{r}
plot_compare(compare_fmb, "FMB_XDC")
```
```{r}
summary_fmb <- summary_statistics(filtered_dfs_fmb$new, filtered_dfs_fmb$old, compare_fmb)
summary_fmb
```

## GG_GEI_G01_XDC 

```{r}
dfs_gg_gei <- get_dfs("GG_GEI_G01_XDC")
visualize(dfs_gg_gei$old, dfs_gg_gei$new, "GG_GEI_G01_XDC")
```
```{r}
get_country_diffs(dfs_gg_gei$old, dfs_gg_gei$new)
```

```{r}
filtered_dfs_gg_gei <- filter_diffs(dfs_gg_gei$old, dfs_gg_gei$new)
compare_gg_gei <- compare_values(filtered_dfs_gg_gei$old, filtered_dfs_gg_gei$new, tolerance = 1e+12)
```

```{r}
summary_gg_gei <- summary_statistics(filtered_dfs_gg_gei$new, filtered_dfs_gg_gei$old, compare_gg_gei)
```

## BCG_GEI_G01_XDC
 
```{r}
dfs_bcg_gei <- get_dfs("BCG_GEI_G01_XDC")
visualize(dfs_bcg_gei$old, dfs_bcg_gei$new, "BCG_GEI_G01_XDC")
```

```{r}
filtered_dfs_bcg_gei <- filter_diffs(dfs_bcg_gei$old, dfs_bcg_gei$new)
compare_bcg_gei <- compare_values(filtered_dfs_bcg_gei$old, filtered_dfs_bcg_gei$new, tolerance = 1e+12)
summary_bcg_gei <- summary_statistics(filtered_dfs_bcg_gei$new, filtered_dfs_bcg_gei$old, compare_bcg_gei)
```

## ENDA_XDC_USD_RATE

quicly reviewing the first one again, just to check if the functions work as expected

```{r}
dfs_xdc <- get_dfs("ENDA_XDC_USD_RATE")
filtered_dfs_xdc <- filter_diffs(dfs_xdc$old, dfs_xdc$new)
compare_xdc <- compare_values(filtered_dfs_xdc$old, filtered_dfs_xdc$new, tolerance = 0.001)
summary_xdc <- summary_statistics(filtered_dfs_xdc$new, filtered_dfs_xdc$old, compare_xdc)
```



## Summaries

```{r}
summary_all <- bind_rows(
  summary_xdc,
  summary_pcpi,
  summary_pcpi_ix,
  summary_bcaxf,
  summary_fmb,
  summary_gg_gei,
  summary_bcg_gei
)
summary_all
```

