# Statistics 

comparing the old and the new one (especially the annual data)

## Loading

```{r}
#| message: false
library(tidyverse)
library(visdat)
library(patchwork)
library(skimr)
```

```{r}
#| message: false
#| warning: false
df_new <- read_csv("../../data/imf_processed/all_annual.csv")

df_old_adv <- read_csv("../../examples/imf/adv_annual.csv")
df_old_dev <- read_csv("../../examples/imf/developing_annual.csv")
df_old_oth <- read_csv("../../examples/imf/other_annual.csv")
```

Combine the old ones
```{r}
df_old_comb <- tibble()
for (df in list(df_old_adv, df_old_dev, df_old_oth)) {
  df <- df %>%
    select(`Country Name`, `Country Code`, `Indicator Name`, `Indicator Code`, Attribute, `1960`:`2024`) %>%
    mutate(across(`1960`:`2024`, as.numeric))
  df_old_comb <- bind_rows(df_old_comb, df) 
}
```

```{r}
df_new <- df_new %>%
  select(`Country Name`, `Country Code`, `Indicator Name`, `Indicator Code`, Attribute, `1960`:`2024`) %>%
  mutate(across(`1960`:`2024`, as.numeric)) %>%
  arrange(`Indicator Code`, `Country Code`)
```

Filtering the old one for the new indicators
```{r}
df_old <- df_old_comb %>%
  filter(
    `Indicator Code` %in% unique(df_new$`Indicator Code`),
    Attribute == "Value"
    ) %>%
  arrange(`Indicator Code`, `Country Code`)
```

## Comparison

```{r}
indicator <- "ENDA_XDC_USD_RATE"
df_new_ind <- df_new %>%
  filter(`Indicator Code` == indicator)
df_old_ind <- df_old %>%
  filter(`Indicator Code` == indicator)
```

Lets start with a visual comparison
```{r}
vis_dat(df_old_ind) + vis_dat(df_new_ind) +
  plot_annotation(title = indicator) +
  plot_layout(guides = "collect")
```
=> Looks very similar in structure

Differences in country codes present in new, but not in old
```{r}
present_new <- setdiff(df_new_ind$`Country Code`, df_old_ind$`Country Code`)
df_new_ind %>%
  filter(`Country Code` %in% present_new) %>%
  select(`Country Name`, `Country Code`) %>%
  distinct()
```
= well USSR is quote outdated, and Caymen Islands is not that important

Other way around
```{r}
present_old <- setdiff(df_old_ind$`Country Code`, df_new_ind$`Country Code`)
df_old_ind %>%
  filter(`Country Code` %in% present_old) %>%
  select(`Country Name`, `Country Code`) %>%
  distinct()
```
= seems to be mostly small islands belinging to the UK, regional groupings, and an old code for North yemen (existed until 1990)

Lets filter them out
```{r}
df_old_ind <- df_old_ind %>% 
  filter(!`Country Code` %in% present_old) %>%
  mutate(`Country Code` = as.character(`Country Code`))
  
df_new_ind <- df_new_ind %>% filter(!`Country Code` %in% present_new)
```

Now we can compare the actual values
```{r}
compare <- diffdf::diffdf(
  df_old_ind, df_new_ind,
  strict_numeric = F,
  keys = c("Country Code"),
  tolerance = 0.001
)
```

```{r}
compare$NumDiff
# compare                # print all of it
# compare$VarDiff_2017   # example of differences in 2017, the old one does not have data for it...
```

- Some names differ for countries (as expected)
- the description has a typo (that should not be relevant)
- Values are almost identical (just in some years there is one or two observations missing)

For quick comparison, lets build some summary statistics 
```{r}
summary_old <- df_old_ind %>%
  pivot_longer(cols = `1960`:`2024`, names_to = "Year", values_to = "Value") %>%
  group_by(Year) %>%
  summarise(
    mean = mean(Value, na.rm = T),
    sd = sd(Value, na.rm = T),
    min = min(Value, na.rm = T),
    max = max(Value, na.rm = T),
    n = n(),
    n_na = sum(is.na(Value))
  ) %>%
  mutate(Source = "Old")

summary_new <- df_new_ind %>%
  pivot_longer(cols = `1960`:`2024`, names_to = "Year", values_to = "Value") %>%
  group_by(Year) %>%
  summarise(
    mean = mean(Value, na.rm = T),
    sd = sd(Value, na.rm = T),
    min = min(Value, na.rm = T),
    max = max(Value, na.rm = T),
    n = n(),
    n_na = sum(is.na(Value))
  ) %>%
  mutate(Source = "New")

summary_comb <- bind_rows(summary_old, summary_new) %>%
  mutate(Year = as.integer(Year))
summary_comb
```

```{r}
skim_df1 <- skim(df_new_ind) %>% reframe(skim_variable, numeric.mean, n_missing,source = "new")
skim_df2 <- skim(df_old_ind) %>% reframe(skim_variable, numeric.mean, n_missing,source = "old")

skim_df1 %>% 
  bind_rows(skim_df2) %>%
  arrange(skim_variable, source)
```

